---
title: "Forecasting Bankruptcy Rates"
author: "Chris Dong"
date: "November 28, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, warning = F, message = F)
```

#### Libaries
```{r, warning = F, message = F}
library(tseries)
library(car)
library(forecast)
library(tidyverse)
library(magrittr)
library(ggcorrplot)
```

### Loading Data
```{r, warning = F, message = F}
train <- read_csv("train.csv")
test <- read_csv("test.csv")

train %<>% na.omit()
bank <-ts(train$Bankruptcy_Rate, start = c(1987, 1), end =c(2010, 12), frequency = 12)
house <-ts(train$House_Price_Index, start = c(1987, 1), 
           end =c(2010, 12), frequency = 12)
unemployment <- ts(train$Unemployment_Rate, start = c(1987, 1), 
           end =c(2010, 12), frequency = 12)
population <- ts(train$Population, start = c(1987, 1), 
           end =c(2010, 12), frequency = 12)
```

### Create Training and Validation Set
```{r}
bank.train <- window(bank, start = c(1987,1), end = c(2008,12))
bank.test <- window(bank, start = c(2009,1), end = c(2010,12))
house.train <- window(house, start = c(1987,1), end = c(2008,12))
house.test <- window(house, start = c(2009,1), end = c(2010,12))
unemployment.train <- window(unemployment, start = c(1987,1), end = c(2008,12))
unemployment.test <- window(unemployment, start = c(2009,1), end = c(2010,12))
population.train <- window(population, start = c(1987,1), end = c(2008,12))
population.test <- window(population, start = c(2009,1), end = c(2010,12))
```

```{r}
plot(log(bank.train))
```

```{r}
adf.test(bank.train)
```

```{r}
bank.train1 <- diff(bank.train)
```

```{r}
adf.test(bank.train1)$p.value
```

```{r}
bank.train2 <- diff(bank.train1, lag = 12)
```

Trying `auto.arima` as baseline
```{r}
automl <- arima(log(bank.train), order = c(2,0,1),
      seasonal = list(order = c(0,0,2), method = "ML"))
sqrt(mean((exp(forecast(automl, level = 95, h = 24)$mean) - bank.test)^2))
```

```{r, eval = F}
result <-c()
orderlist = list()

for(i in 0:3){
  for(j in 0:3){
    for(a in 0:3){
      for(b in 0:3){
        
        orderlist <- c(orderlist, paste(i,j,a,b))
        bankmodel <- tryCatch({expr = arima(log(bank.train), order = c(i,1,j), 
          seasonal = list(order = c(a,1,b), period = 12), method = "ML")},
          error = function(cond) {return(NA)})
        
        rmse <- sqrt(mean((exp(forecast(bankmodel, 
                level = 95, h = 24)$mean) - bank.test)^2))
        print(paste(i,j,a,b, ":", rmse))
        ifelse(!is.na(bankmodel), result <- c(result, rmse), 
               result <- c(result, NA))
  }
  }
  }
}
save(result, orderlist, file = "bank.RData")
```

```{r}
load(file = "bank.RData")
```


```{r}
head(result[order(result)], n = 25)
```

```{r}
m1 <- arima(log(bank.train), order = c(0,1,3), seasonal = list(order = c(2,1,3), period = 12), method = "ML")
m2 <- arima(log(bank.train), order = c(0,1,3), seasonal = list(order = c(3,1,3), period = 12), method = "ML")
D <- -2*(m1$loglik - m2$loglik)
pval <- 1-pchisq(D,length(m2$coef) - length(m1$coef))
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```

SARIMA(0,1,3)(2,1,3) better than SARIMA(0,1,3)(3,1,3)

```{r}
m1 <- arima(log(bank.train), order = c(0,1,2), seasonal = list(order = c(2,1,3), period = 12), method = "ML")
m2 <- arima(log(bank.train), order = c(0,1,3), seasonal = list(order = c(2,1,3), period = 12), method = "ML")
D <- -2*(m1$loglik - m2$loglik)
pval <- 1-pchisq(D,length(m2$coef) - length(m1$coef))
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```

SARIMA(0,1,3)(2,1,3) better than SARIMA(0,1,2)(2,1,3)

```{r}
rmse <- function(logmodel) sqrt(mean((exp(forecast(logmodel, level = 95, h = 24)$mean) - bank.test)^2))
```

```{r}
model <-  arima(log(bank.train), order = c(0,1,3), seasonal = list(order = c(2,1,3), period = 12), method = "ML")
(score <- rmse(model))
```

So far, an SARIMA(0,1,3)(2,1,3) gets a RMSE of `r score` when forecasting from January 2009 to December 2010.

## Holt-Winters

```{r}
hw <- HoltWinters(x = log(bank.train),seasonal = "add",
        alpha = 0.2, beta = 0.2, gamma = 0.4) 
rmse(hw)
```

## Additive Triple Exponential Smoothing
```{r, eval = F}
hwresult <-c()
hworderlist = list()

for(i in seq(0.001,1, by = 0.001)){
  for(j in seq(0.001,1, by = 0.001)){
    for(a in seq(0.001,1, by = 0.001)){
        
        hworderlist <- c(hworderlist, paste(i,j,a))
        bankmodel <- HoltWinters(x = log(bank.train),seasonal = "add",
                     alpha = i, beta = j, gamma = a) 
        measure <- rmse(bankmodel)
        print(paste(i,j,a, ":", measure))
        hwresult <- c(hwresult, measure)
  
  }
  }
}
save(hwresult, hworderlist, file = "holt.RData")
```


```{r}
head(holt_add[order(holt_add)], n = 10)
```

## Multiplicative Triple Exponential Smoothing
```{r, eval = F}
hwresult <-c()
hworderlist = list()

for(i in seq(0.05,1, by = 0.05)){
  for(j in seq(0.05,1, by = 0.05)){
    for(a in seq(0.05,1, by = 0.05)){
        
        hworderlist <- c(hworderlist, paste(i,j,a))
        bankmodel <- HoltWinters(x = log(bank.train),seasonal = "mult",
                     alpha = i, beta = j, gamma = a) 
        measure <- rmse(bankmodel)
        print(paste(i,j,a, ":", measure))
        hwresult <- c(hwresult, measure)
  
  }
  }
}
save(hwresult, hworderlist, file = "holt2.RData")
```

```{r}
head(holt_mult[order(holt_mult)], n = 10)
```

```{r}
#save(result, holt_add, holt_mult, file = "bank.RData")
```

Multivariate

#### Correlation Matrix
```{r}
train[,-1] %>% na.omit() %>% cor() %>% ggcorrplot(lab = T)
```

# SARIMAX
```{r}
model.population <-  arima(log(bank.train), order = c(0,1,3), 
                seasonal = list(order = c(2,1,3), period = 12), 
                method = "ML",
                xreg = data.frame(population.train))
```

```{r}
(score2 <- sqrt(mean((exp(forecast(model.population, level = 95, h = 24,
                xreg = data.frame(population.test))$mean) - bank.test)^2)))
```

Population improved RSME from `r score` to `r score2`.

### Trying Population + Housing Price Index
```{r}
model.unemploy.pop <-  arima(log(bank.train), order = c(0,1,3), 
                seasonal = list(order = c(2,1,3), period = 12), 
                method = "ML",
                xreg = data.frame(population.train, house.train))
```

```{r}
(score3 <- sqrt(mean((exp(forecast(model.unemploy.pop, level = 95, h = 24,
                xreg = data.frame(population.test, house.test))$mean) - bank.test)^2)))
```

## Comparing Population and (Housing + Population) with Log-Likelihood Test

```{r}
D <- -2*(model.population$loglik - model.unemploy.pop$loglik)
pval <- 1-pchisq(D,length(model.unemploy.pop$coef) - length(model.population$coef))
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```

Having both variables is indeed better.

### Trying three variables

```{r}
model.allthree <-  arima(log(bank.train), order = c(0,1,3), 
                seasonal = list(order = c(2,1,3), period = 12), 
                method = "ML",
                xreg = data.frame(population.train, house.train, unemployment.train))
```

```{r}
(score4 <- sqrt(mean((exp(forecast(model.allthree, level = 95, h = 24,
                xreg = data.frame(population.test, house.test,
                                  unemployment.test))$mean) - bank.test)^2)))
```

Doesn't seem better, let's try running a log-likelihood test

```{r}
D <- -2*(model.unemploy.pop$loglik - model.allthree$loglik)
pval <- 1-pchisq(D,length(model.allthree$coef) - length(model.unemploy.pop$coef))
print(c("Test Statistic:",round(D, 4),"P-value:", round(pval, 4)))
```

Taking the log of Population
```{r}
model.unemploy.pop.log <-  arima(log(bank.train), order = c(0,1,3), 
                seasonal = list(order = c(2,1,3), period = 12), 
                method = "ML",
                xreg = data.frame(log(population.train), house.train))
(score3.log <- sqrt(mean((exp(forecast(model.unemploy.pop.log, level = 95, h = 24,
                xreg = data.frame(log(population.test), house.test))$mean) - bank.test)^2)))
```

```{r}
model.unemploy.pop.log2 <-  arima(log(bank.train), order = c(0,1,3), 
                seasonal = list(order = c(2,1,3), period = 12), 
                method = "ML",
                xreg = data.frame(log(population.train), log(house.train)))
(score3.log2 <- sqrt(mean((exp(forecast(model.unemploy.pop.log2, level = 95, h = 24,
                xreg = data.frame(log(population.test), log(house.test)))$mean) - bank.test)^2)))
```

Why is this so low..? Any Ideas?


# Current Best Model

**The best model so far is a SARIMAX (0,1,3)(2,1,3) along with the explanatory variables Population and Housing Price Index. It has a RMSE of `r score3.log`.**